---
title: "Final Project"
author: "Max Cohen"
date: "12/11/2019"
output: ioslides_presentation
---

##The code examined here is a collection of various readings from clam habitats in Canada and the islands of the Puget Sound. These readings and their analysis are from "Spatial and Temporal Coherence in Intertidal Clams". The study investigated how different factors such as salinity and air and sea temperature affect biomass in clam populations over a period of several decades. 

##Here are the programs in R that I used
``` {r} 
## Required Packages 
library(tidyverse) 
library(dplyr) 
library(ggplot2) 
``` 

## Now we are reading in the data files
```{r}
clam_general_data <- read.csv("data/SuppTable1_ClamData_DFA.csv")
clam_salinity_data <- read.csv("data/SuppTable2_Salinity_DFA.csv") 
clam_temp_data <- read.csv("data/SuppTable3_AirTemp_DFA.csv")  
rawseatemp_data <- read.csv("data/SuppTable5_SST_DFA.csv")
```

##Here are the data transformations I did
```{r}
## Here we are averaging each data column for the general clam biomass data
mean_clam_general_data <- mutate(clam_general_data, Mean = rowMeans(select(clam_general_data, starts_with("sa")), na.rm = TRUE)) 
print(mean_clam_general_data)
```

```{r}
## Filters the data to separate values with salinity > 31
filter(clam_salinity_data, salinity > 31) 
## We can see here that salinity was lowest in the 90's, and started to increase thereafter
```

```{r}
## Averages the three separate columns for sea surface temperature reading in to one column sorted by year 
edited_seatemp_data <- mutate(rawseatemp_data, Mean = rowMeans(select(rawseatemp_data, starts_with("sst")), na.rm = TRUE)) 
print(edited_seatemp_data)
```

```{r}
## This code joins the clam_salinity_data table with the table for air temperature. Doing this lets us see the contrasting patterns in temperature and salinity side by side. 
inner_join(clam_temp_data, clam_salinity_data)
```

## Here are the graphs I made
```{r}
## This code graphs the salinity data
ggplot(data = clam_salinity_data) +
  geom_point(mapping = aes(x = year, y = salinity)) + 
  labs(y = "salinity", title = "Salinity Readings Over Time")
## We see from this plot that generally, salinity is decreasing over time. 
```

```{r}
## This code graphs the air temperature data 
ggplot(data = clam_temp_data) +
  geom_point(mapping = aes(x = year, y = airtemp)) + 
  labs(y = "air temperature", title = "Air Temperature Readings Over Time")
##While air temperature readings are varied for the first 20-25 years, after 2010 they begin to climb rapidly. This is consistent with global air temperatures rising as well. 
```

```{r}
##This code graphs the data for sea surface temperature  
ggplot(data = edited_seatemp_data) +
  geom_point(mapping = aes(x = year, y = Mean)) +
  labs(y = "surface temperature", title = "Mean Sea Surface Temperature Over Time") 
##Surface Temperature appears to be becoming more increasingly variable, which is in line with what we know about global climate change
```

```{r}
##Finally we are graphing the averages for the clam biomass data 
ggplot(data = mean_clam_general_data) +
  geom_point(mapping = aes(x = year, y = Mean)) + 
  labs(y = "biomass", title = "Average Clam Biomass Per Year") 
##Biomass readings went up since their lowest point in 1990. This could reflect a number of things, including a decreased rate of global climate change since the late 20th century, and an increase in environmental legislation and protection
```

## In Conclusion, this investigation has revealed that patterns of global change are affecting clams and other shellfish. While the biomass readings do seem to refute this, it would be more than likely that that is due to a number of non-climate-related factors